---
title: "Fairtrade consumption and social milieu"
author: "Philipp D. P."
date: "30.09.2020"
output:
  md_document:
    variant: markdown_github
---
#### Data: Umweltbewusstsein in Deutschland 2016 (Welle 1) - Version 1.0.0 (12.04.2017)
#### Data access: https://doi.org/10.4232/1.12764 (Gesis-Archiv - SPSS Data - SAV Format)
#### R Version 4.0.2 (22.06.2020) - RStudio Version 1.3.1056
<br> 

#### Clear workspace
```{r}
rm(list = ls())
```


```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
Packages <- c("tidyverse",     ## 1.3.0
            "naniar",        ## 0.5.0
            "haven",         ## 2.2.0
            "survey",        ## 3.37
            "car",           ## 3.0-6
            "gmodels",       ## 2.18.1
            "psych",         ## 1.9.12.31
            "mfx",           ## 1.2-2
            "epiDisplay",    ## 3.5.0.1
            "effects",       ## 4.1-4
            "DescTools",     ## 0.99.37
            "margins"        ## 0.3.23
            )

lapply(Packages, require, character.only = T)
```

#### Load packages
```{r eval=F, message=FALSE}
Packages <- c("tidyverse",     ## 1.3.0
            "naniar",        ## 0.5.0
            "haven",         ## 2.2.0
            "survey",        ## 3.37
            "car",           ## 3.0-6
            "gmodels",       ## 2.18.1
            "psych",         ## 1.9.12.31
            "mfx",           ## 1.2-2
            "epiDisplay",    ## 3.5.0.1
            "effects",       ## 4.1-4
            "DescTools",     ## 0.99.37
            "margins"        ## 0.3.23
            )

lapply(Packages, require, character.only = T)
```


### Data import and preparation (preparation of the analysis)

#### Read in data set
```{r}
Data <- read_sav("Umweltbewusstsein_DE_2016_Welle1_v1-0-0.sav")
```

#### Create variable subset 
```{r}
Data <- Data %>%
  dplyr::select(f624_3, ## If possible, only buy products that have been produced under fair working conditions.
                f821,   ## gender (m = 1 / w = 2) 
                f8210,  ## Total monthly net income of a household (in 11 income groups).
                nf822,  ## Age of respondents in years.
                f823,   ## Educational attainment (highest school degree or university degree). 
                QCL_1,  ## Social milieus (six social segments, 17 items for identification).
                weight  ## Weighting variable (weighting by region, gender, age and education).
                )
```

#### Define missing data or unusable values as "NA"
```{r}
Data <- Data %>%
  mutate(f624_3 = na_if(f624_3, 5), ## "don't know"
         f8210  = na_if(f8210, 12), ## "not specified"
         f823   = na_if(f823, 1),   ## "still student"
         f823   = na_if(f823, 7),   ## "Other school diploma"
         f823   = na_if(f823, 8)    ## "not specified"
         )
```

###  Recode / dichotomize variables for analysis
#### Dichotomize target variable purchase probability as new variable "buy"
#### Recode categories 1 and 2 to 1 = Yes / categories 3 and 4 to 0 = No
```{r}
Data <- Data %>%
  mutate(buy = car::recode(Data$f624_3,
                            "1:2 = 1; 3:4 = 0; else = NA", as.factor = T))
```

#### Gender from m = 1 / w = 2 to m = 0 / w = 1
#### New variable "female" (female = 1)
```{r}
Data$female <- dummy.code(Data$f821, group = 2)
```

#### Define milieu as factor variable and thus automatically create a dummy variable
#### New variable "milieu" (milieu 1 automatically set as reference category)
#### Means: The first milieu is included in the calculation with 0 and the other categories with 1 each.
```{r}
Data <- Data %>%
  mutate(milieu = car::recode(Data$QCL_1,
                            "1 = 1;
                             2 = 2;
                             3 = 3;
                             4 = 4; 
                             5 = 5;
                             6 = 6", 
                             as.factor = T))
```

#### Distribution of milieus rounded to one decimal place
```{r}
round(prop.table(table(Data$milieu))*100, digits = 1)
```

#### How exactly is the variable "milieu" structured as a factor. Important for later model interpretation.
```{r}
contrasts(Data$milieu)
```

#### Change reference category of milieu variable to 4 (preparation for logistic regression).
```{r}
Data$milieu <- relevel (Data$milieu, ref = "4")
```

#### Structure of the variable after changing the reference category
#### Fourth milieu (precarious) set as reference and assigned value 0
```{r}
contrasts(Data$milieu)
```

#### Define education as factor variable and thus automatically create a dummy variable
#### New variable "education" ("without degree" automatically set as reference category)
```{r}
Data$education <- factor(Data$f823)
```

#### Add age variable as "age" for clarity
```{r}
Data$age <- Data$nf822
```

#### Add income variable as "income" for clarity
```{r}
Data$income <- Data$f8210
```

#### Missing values analysis 

```{r}
summary(Data)
```

#### Visualization of missing values
```{r, fig.width=10,fig.height=6}
variables <-data.frame(Data$female, Data$education, Data$income, Data$age, Data$milieu, Data$buy)
gg_miss_var(variables)
```

### Check application requirements of binary logistic regression.

#### 1. No extreme unequal distribution of the dependent variable.
```{r}
prop.table(table(Data$buy))*100
```

```{r include=FALSE}
## Regression vor Berechnung der vif-Werte, damit Berechnung möglich wird
m6.w <- glm(buy ~ female + education + income + age + milieu,
            family = binomial,
            data = Data,
            weights = weight)
summary(m6.w)
```

#### 2. Sufficient cases in predictor variable categories associated with outcome variable.
```{r}
CrossTable(Data$milieu, Data$buy, prop.r=T, prop.c=F, prop.t=F, prop.chisq=F, digits = 2) 
```

#### 3. Low multicollinearity (Checked for overall model).
#### variance inflation factor (VIF) (should not exceed 10).
```{r}
vif(m6.w)
```

### Binary logistic regression model

#### Dependent variable: buy from Fairtrade Yes (1) / No (0)
#### Independent variables: Gender, education, income, age, milieu
#### Exclusion of cases with missing values ("na.action=na.omit" set as default)
#### Models (each separately and at the end together in one model)
#### Error message "non-integer #successes in a binomial glm!" comes up in each case.
#### This is due to the weighting, since there are no integer values left
#### The error message is therefore hidden in the following output 
<br>

#### Zero model (unweighted)
```{r}
m.null <- glm(buy ~ 1,
              family = binomial,
              data = Data)
summary(m.null)
```

#### gender (weighted)
```{r warning=FALSE}
m1.w <- glm(buy ~ female,
            family = binomial,
            data = Data,
            weights = weight)
summary(m1.w)
```

#### education (weighted)
```{r warning=FALSE}
m2.w <- glm(buy ~ education,
            family = binomial,
            data = Data,
            weights = weight)
summary(m2.w)
```

#### income (weighted)
```{r warning=FALSE}
m3.w <- glm(buy ~ income,
            family = binomial,
            data = Data,
            weights = weight)
summary(m3.w)
```

#### age (weighted)
```{r warning=FALSE}
m4.w <- glm(buy ~ age,
            family = binomial,
            data = Data,
            weights = weight)
summary(m4.w)
```

#### milieu (weighted)
```{r warning=FALSE}
m5.w <- glm(buy ~ milieu,
            family = binomial,
            data = Data,
            weights = weight)
summary(m5.w)
```

#### Overall model (weighted) - socio-demographics as control variables.
```{r warning=FALSE}
m6.w <- glm(buy ~ female + education + income + age + milieu,
            family = binomial,
            data = Data,
            weights = weight)
summary(m6.w)
```

#### Model comparison and model quality
#### Log-Likelihood
```{r}
logLik(m.null)
logLik(m5.w) 
logLik(m6.w)
```

#### Likelihood-Ratio-Test
```{r}
lrtest(m.null, m5.w)
```

#### AIC - Akaike Information Criterion
```{r warning=FALSE}
AIC(m.null, m5.w, m6.w)
```


#### BIC - Bayesian Information Criterion 
```{r warning=FALSE}
BIC(m.null, m5.w, m6.w)
```

#### Pseudo R2 (McKelvey & Zavoina)
#### Directly in percent and rounded to one decimal place
```{r echo=TRUE, warning=FALSE}
round((PseudoR2(m1.w, "McKelveyZavoina")*100), digits = 1)   ## gender
round((PseudoR2(m2.w, "McKelveyZavoina")*100), digits = 1)   ## education
round((PseudoR2(m3.w, "McKelveyZavoina")*100), digits = 2)   ## income
round((PseudoR2(m4.w, "McKelveyZavoina")*100), digits = 1)   ## age
round((PseudoR2(m5.w, "McKelveyZavoina")*100), digits = 1)   ## milieu
```

#### Overall model - R2 (McKelvey & Zavoina)

```{r warning=FALSE}
round((PseudoR2(m6.w, "McKelveyZavoina")*100), digits = 1) ## Overall model 
```

#### Proportion of explained variance by milieu in total explained variance.
```{r}
round(11.9/21.2*100, digits = 1)
```

#### Advanced analysis of the models

#### Odds Ratio
```{r warning=FALSE}
logistic.display(m5.w)
logistic.display(m6.w)
```

#### AME (Average Marginal Effects)
```{r}
m5.w.marginal <- margins(m5.w)
summary(m5.w.marginal)

m6.w.marginal <- margins(m6.w) 
summary(m6.w.marginal)
```

#### Visualization of the AME
```{r, fig.width=10,fig.height=6}
m5.w.marginal.sum <- summary(m5.w.marginal)

ggplot(data = m5.w.marginal.sum) +
  geom_point(aes(factor, AME)) +
  geom_errorbar(aes(x = factor, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

```{r, fig.width=10,fig.height=6}
m6.w.marginal.sum <- summary(m6.w.marginal)

ggplot(data = m6.w.marginal.sum) +
  geom_point(aes(factor, AME)) +
  geom_errorbar(aes(x = factor, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

